Microxanox: an R package for simulating an aquatic \(MICR\)obial ecosystem that can occupy \(OX\)ic or \(ANOX\)ic states.

### 2022-06-16 inst/manuscript/manuscript_microxanox.Rmd
Final draft

---
title: "Microxanox: an R package for simulating an aquatic $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch;  Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
keywords: 
  - reproducibility
  - regime shift 
  - stable state
  - ordinary differential equations
abstract: |
  $Microxanox$ is an R package to simulate a three functional group system (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) osing a set of ordinary differential equations. Simulations can be run individually or over a parameter range, to find stable states.  The model features multiple species per functional group, where the number is only limited by computational constraints.
  The R package is constructed in such a way, that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated.
journal: "SoftwareX"
date: "`r Sys.Date()`"

@@

| **Nr.** | **Code metadata description**                                | **Please fill in this column**                                                 |  
| :-----: | :----------------------------------------------------------- | :----------------------------------------------------------------------------- |  
|   C1    | Current code version                                         | v0.9.0                                                                         |  
|   C2    | Permanent link to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) | 
|         |                                                              | [10.5281/zenodo.6624125](https://zenodo.org/record/6624125)                    |
|   C3    | Code Ocean compute capsule                                   |                                                                                |  
|   C4    | Legal Code License                                           | CC BY 4.0                                                                      |  
|   C5    | Code versioning system used                                  | git                                                                            |  
|   C6    | Software code languages, tools, and services used            | [R](https://cran.r-project.org/index.html)                                     |  
|   C7    | Compilation requirements, operating environments             | [R (>= 4.1.0)](https://cran.r-project.org/index.html)                          |  
|         |                                                              | magrittr                                                                       |  
|         |                                                              | tibble                                                                         |  
|         |                                                              | ggplot2                                                                        |  
|         |                                                              | patchwork                                                                      |  
|         |                                                              | grDevices                                                                      |  
|         |                                                              | stats                                                                          |  
|         |                                                              | mgcv                                                                           |  
|         |                                                              | deSolve                                                                        |  
|         |                                                              | dplyr                                                                          |  
|         |                                                              | tidyr                                                                          |  
|         |                                                              | stringr                                                                        |  
|         |                                                              | multidplyr                                                                     |  
|   C8    | If available Link to developer documentation/manual          | [Usert Manual](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) |  
|   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Rainer@krugs.de                                            |

@@

The framework used when writing this package aimed to maximise simplicity for the user, and to make it easy to reproduce results (see the supplement to @Limberger2022 for an example of how this is used). As such, all the parameters needed to run a simulation or find a stable state (i.e. the final state of the ecosystem) are contained in a single object (which can easily be created using included functions). This parameter object is given to a function that runs the simulations and returns the results. The returned results object is identical to the parameter object but with an additional slot named `results`, which contains the results of the run. Thus the returned results object contains the simulation conditions (parameters) as well as the results, and can be used to run the simulation again. This promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations straightforward.

@@

A typical simulation would look as shown in \ref{fig:runsim_example}.

@@

The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function. This function is only provided as a convenience function to provide a way to easily see the results of a simulation run. An example plot resulting from this function is shown in \ref{fig:plot-dynamics}.

```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap="Results of a simulation run shown using the function $plot_dynamics()$. In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example. CB\textunderscore1 = cyanobacteria strain 1; SB\textunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show the figure produce by $plot_dynamics()$, which produces figures not with all features ideal for publication.", fig.align="center", out.width = 350 }
knitr::include_graphics("figures/ug_three_strains_dynamics.png")
```

@@

Here we evidence the impact of the *microxanox* package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems. The first two use cases are described in detail (including the code for reproducing them) in the *User Guide* and the *Partial Reproduction* vignettes. The third is taken from @REF_NEEDED.

@@

An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \ref{fig:uc1}. The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria. Oyxgen diffusivity is then slowly decreased and eventually, around hour 30'000 the system switches to the anoxic state, with high abundance of both sulfur bacteria types. The oxygen diffusivity is then increased and at around hour 38'000 the system abruptly switches back to the oxic state.

@@

```{r uc1, echo = FALSE,  include=TRUE, fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes. Plot of the stable states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB\textunderscore1 = cyanobacteria strain 1; SB\textunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot_dynamics()$  function.', fig.align="center", out.width = 350}
fig <- here::here("inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching.pdf")
#if (!file.exists(fig)){
  library(ggplot2)
  library(microxanox)
  library(dplyr)
  x <- readRDS(here::here("inst/manuscript/data/hlh_with_defaults.RDS"))
  p <- plot_dynamics(x)
  p[[1]] <- p[[1]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[2]] <- p[[2]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[3]] <- p[[3]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p_a <- p[[4]]
  p[[4]]$data <- p[[4]]$data %>% dplyr::filter(species != "a")
  p[[4]]$data$species <- p[[4]]$data$species %>%
    dplyr::recode(SO = "Oxidised sulfur", O = "Oxygen", SR = "Sulfide", P = "Phosphorus")
  p[[4]] <- p[[4]] +
    xlab("time (h)") + 
    ylab(expression(atop(log[10](concentration),(µM)))) + 
    labs(colour = "substrate") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p_a$data <- p_a$data %>% dplyr::filter(species == "a")
  p_a$data$species <- p_a$data$species %>%
    dplyr::recode(a = "Oxygen diffusivity ")
  p_a <- p_a +
    ylab(expression(atop(Oxygen~diffusivity,(hr^{-1})))) + 
    labs(colour = "Environmental driver") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p <- p[[1]] / p[[2]] / p[[3]] / p[[4]] / p_a
  ggsave(
    filename = fig,
    plot = p,
    device = "pdf",
    width = 8,
    height = 8
  )  
#}
knitr::include_graphics(fig)
```

@@

The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \ref(figL:uc2)). For example, with only the CB (cyanobacteria) present, there was no hysteresis. In contrast, the presence of both CB and SB (sulfate reducing bacteria) led to a large amount of hysteresis. (These results are also given in the *Partial Reproduction* vignette.)

```{r uc2, echo = FALSE,  include=TRUE, fig.cap='The amount of hysteresis depends on the combination of types of organisms present. The model is entirely deterministic, hence there are no error bars.', fig.align="center", out.width = 150 }
fig <- here::here("inst/manuscript/figures/gen_uc2_user_guide_hysteresis.pdf")
#if (!file.exists(fig)){
  library(ggplot2)
  x <- readRDS(here::here("inst/manuscript/data/resp_summ5.rds"))
  x$composition <- as.character(x$composition)
  x <- x[order(x$mean_mean_hyst),]
  x$composition <- factor(x$composition, x$composition, ordered = TRUE)

@@

 ggsave(
   filename = fig,
   plot = p,
   device = "pdf",
   width = 3, height = 3
 )

#}
knitr::include_graphics(fig)
```

@@

The *microxanox* R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present. It has been used for the research described in another paper that describes one of the first investigations of the effects of diversity on ecosystem resilience @Limberger2022. In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity. The *microxanox* package has also been used to reproduce the results of the paper that inspired the package development @Bush2017.

>>>>

---
title: "Microxanox: an R package for simulating an aquatic $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states."
author:
- name: Rainer M Krug
  email: Rainer.Krug@uzh.ch;  Rainer@krugs.de
  affiliation: University of Zürich
  footnote: Corresponding Author
- name: Owen L. Petchey
  email: Owen.Petchey@ieu.uzh.ch
  affiliation: University of Zürich
address:
- code: University of Zürich
  address: Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse
    190, 8057 Zurich
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
keywords: 
  - reproducibility
  - regime shift 
  - stable state
  - ordinary differential equations
abstract: |
  $Microxanox$ is an R package to simulate a three functional group system (cyanobacteria, phototrophic sulfur bacteria, and sulfate-reducing bacteria) with four chemical substrates (phosphorus, oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary differential equations. Simulations can be run individually or over a parameter range, to find stable states.  The model features multiple species per functional group, where the number is only limited by computational constraints.
  The R package is constructed in such a way, that the results contain the input parameter used, so that a saved results can be loaded again and the simulation be repeated.
journal: "SoftwareX"
date: "`r Sys.Date()`"

@@

| **Nr.** | **Code metadata description**                                | **Please fill in this column**                                                 |  
| :-----: | :----------------------------------------------------------- | :----------------------------------------------------------------------------- |  
|   C1    | Current code version                                         | v0.9.0                                                                         |  
|   C2    | Permanent link to code/repository used for this code version | [https://github.com/UZH-PEG/microxanox](https://github.com/UZH-PEG/microxanox) | 
|         |                                                              | [10.5281/zenodo.6624125](https://zenodo.org/record/6624125)                    |
|   C3    | Code Ocean compute capsule                                   |                                                                                |  
|   C4    | Legal Code License                                           | MIT                                                                            |  
|   C5    | Code versioning system used                                  | git                                                                            |  
|   C6    | Software code languages, tools, and services used            | [R](https://cran.r-project.org/index.html)                                     |  
|   C7    | Compilation requirements, operating environments             | [R (>= 4.1.0)](https://cran.r-project.org/index.html)                          |  
|         |                                                              | magrittr                                                                       |  
|         |                                                              | tibble                                                                         |  
|         |                                                              | ggplot2                                                                        |  
|         |                                                              | patchwork                                                                      |  
|         |                                                              | grDevices                                                                      |  
|         |                                                              | stats                                                                          |  
|         |                                                              | mgcv                                                                           |  
|         |                                                              | deSolve                                                                        |  
|         |                                                              | dplyr                                                                          |  
|         |                                                              | tidyr                                                                          |  
|         |                                                              | stringr                                                                        |  
|         |                                                              | multidplyr                                                                     |  
|   C8    | If available Link to developer documentation/manual          | [Usert Manual](https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html) |  
|   C9    | Support email for questions                                  | Rainer.Krug@uzh.ch; Owen.Petchey@ieu.uzh                                       |

@@

The framework used when writing this package aimed to maximise simplicity for the user, and to make it easy to reproduce results (see the supplement [@Petchey2022] to @Limberger2022 for an example of how this is used). As such, all the parameters needed to run a simulation or find a stable state (i.e. the final state of the ecosystem) are contained in a single object (which can easily be created using included functions). This parameter object is given to a function that runs the simulations and returns the results. The returned results object is identical to the parameter object but with an additional slot named `results`, which contains the results of the run. Thus the returned results object contains the simulation conditions (parameters) as well as the results, and can be used to run the simulation again. This promotes reproducibility and makes incremental changes of individual parameters and re-running the simulations straightforward.

@@

A typical simulation would look as shown in Figure \ref{fig:runsim_example}.

@@

The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function. This function is only provided as a convenience function to provide a way to easily see the results of a simulation run. An example plot resulting from this function is shown in Figure \ref{fig:plot-dynamics}.

```{r plot-dynamics, echo = FALSE,  include=TRUE, fig.cap="Results of a simulation run shown using the function $plot\\_dynamics()$. In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example. CB\\_1 = cyanobacteria strain 1; SB\\_1 = sulfur reducing bacteria stain 1; PB\\_1 = phototrophic sulfur bacteria strain 1.", fig.align="center", out.width = 350 }
knitr::include_graphics("figures/ug_three_strains_dynamics.png")
```

@@

Here we evidence the impact of the *microxanox* package by describing three use cases and then by describing how the package can be a starting point for models of other ecosystems. The first two use cases are described in detail (including the code for reproducing them) in the *User Guide* and the *Partial Reproduction* vignettes. The third is taken from @Limberger2022 and @Petchey2022.

@@

An example of this functionality is given in the *Partial Reproduction* vignette, which we briefly show here (Figure \ref{fig:uc1}). The example is composed of a single simulation, at the beginning of which the system is in the oxic state with high abundance of cyanobacteria. Oyxgen diffusivity is then slowly decreased and eventually, around hour 30'000 the system switches to the anoxic state, with high abundance of both sulfur bacteria types. The oxygen diffusivity is then increased and at around hour 38'000 the system abruptly switches back to the oxic state.

@@

```{r uc1, echo=FALSE, fig.align="center", fig.cap='The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the stable states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB\textunderscore1 = cyanobacteria strain 1; SB\\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\\_dynamics()$  function.', message=FALSE, warning=FALSE, include=TRUE, messages=FALSE, out.width=350}
fig <- here::here("inst/manuscript/figures/gen_uc1_partrep_temporal_state_switching.pdf")
if (!file.exists(fig)){
  library(ggplot2)
  library(microxanox)
  library(dplyr)
  x <- readRDS(here::here("inst/manuscript/data/hlh_with_defaults.RDS"))
  p <- plot_dynamics(x)
  p[[1]] <- p[[1]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[2]] <- p[[2]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p[[3]] <- p[[3]] + 
    xlab(NULL) + 
    ylab(expression(atop(log[10](density),(cells~L^{-1})))) +
    theme_bw()
  p_a <- p[[4]]
  p[[4]]$data <- p[[4]]$data %>% dplyr::filter(species != "a")
  p[[4]]$data$species <- p[[4]]$data$species %>%
    dplyr::recode(SO = "Oxidised sulfur", O = "Oxygen", SR = "Sulfide", P = "Phosphorus")
  p[[4]] <- p[[4]] +
    xlab("time (h)") + 
    ylab(expression(atop(log[10](concentration),(µM)))) + 
    labs(colour = "substrate") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p_a$data <- p_a$data %>% dplyr::filter(species == "a")
  p_a$data$species <- p_a$data$species %>%
    dplyr::recode(a = "Oxygen diffusivity ")
  p_a <- p_a +
    ylab(expression(atop(Oxygen~diffusivity,(hr^{-1})))) + 
    labs(colour = "Environmental driver") +
    theme_bw() +
    xlab(expression(time~(hours)))
  p <- p[[1]] / p[[2]] / p[[3]] / p[[4]] / p_a
  ggsave(
    filename = fig,
    plot = p,
    device = "pdf",
    width = 8,
    height = 8
  )  
}
knitr::include_graphics(fig)
```

@@

The results show that the amount of hysteresis depends greatly on the combinations of organisms present (Figure \ref{fig:uc2}). For example, with only the CB (cyanobacteria) present, there was no hysteresis. In contrast, the presence of both CB and SB (sulfate reducing bacteria) led to a large amount of hysteresis. (These results are also given in the *Partial Reproduction* vignette.)

```{r uc2, echo = FALSE,  include=TRUE, fig.cap='The amount of hysteresis depends on the combination of types of organisms present. The model is entirely deterministic, hence there are no error bars.', fig.align="center", out.width = 150 }
fig <- here::here("inst/manuscript/figures/gen_uc2_user_guide_hysteresis.pdf")
if (!file.exists(fig)){
  library(ggplot2)
  x <- readRDS(here::here("inst/manuscript/data/resp_summ5.rds"))
  x$composition <- as.character(x$composition)
  x <- x[order(x$mean_mean_hyst),]
  x$composition <- factor(x$composition, x$composition, ordered = TRUE)

@@

 ggsave(
   filename = fig,
   plot = p,
   device = "pdf",
   width = 3, height = 3
 )  
}
knitr::include_graphics(fig)
```

@@

The *microxanox* R package allows the simulation, visualisation, and analysis of a model of a microbial ecosystem while allowing variation in the amount of diversity containing in each of the functional groups of organisms present. It has been used for the research described in another paper that describes one of the first investigations of the effects of diversity on ecosystem resilience @Limberger2022. In that paper, we show that diversity can have large and important effects of ecosystem responses, highlighting the need for models such as ours, with which one can easily manipulate the amount of biodiversity. The *microxanox* package has also been used to reproduce the results of the paper that inspired the package development [@Bush2017].

### 2022-06-16 inst/manuscript/manuscript_microxanox.tex
Final draft

  \begin{abstract}
  \(Microxanox\) is an R package to simulate a three functional group
  system (cyanobacteria, phototrophic sulfur bacteria, and
  sulfate-reducing bacteria) with four chemical substrates (phosphorus,
  oxygen, reduced sulfur, and oxidized sulfur) osing a set of ordinary
  differential equations. Simulations can be run individually or over a
  parameter range, to find stable states. The model features multiple
  species per functional group, where the number is only limited by
  computational constraints. The R package is constructed in such a way,
  that the results contain the input parameter used, so that a saved
  results can be loaded again and the simulation be repeated.
  \end{abstract}
    \begin{keyword}
    reproducibility \sep regime shift \sep stable state \sep 
    ordinary differential equations
  \end{keyword}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule()
\endhead
C1 & Current code version & v0.9.0 \\
C2 & Permanent link to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
& & \href{https://zenodo.org/record/6624125}{10.5281/zenodo.6624125} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & CC BY 4.0 \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used &
\href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments &
\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
C8 & If available Link to developer documentation/manual &
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{Usert
Manual} \\
C9 & Support email for questions & Rainer.Krug@uzh.ch;
Rainer@krugs.de \\
\bottomrule()
\end{longtable}

@@

The framework used when writing this package aimed to maximise
simplicity for the user, and to make it easy to reproduce results (see
the supplement to \citet{Limberger2022} for an example of how this is
used). As such, all the parameters needed to run a simulation or find a
stable state (i.e.~the final state of the ecosystem) are contained in a
single object (which can easily be created using included functions).
This parameter object is given to a function that runs the simulations
and returns the results. The returned results object is identical to the
parameter object but with an additional slot named \texttt{results},
which contains the results of the run. Thus the returned results object
contains the simulation conditions (parameters) as well as the results,
and can be used to run the simulation again. This promotes
reproducibility and makes incremental changes of individual parameters
and re-running the simulations straightforward.

@@

The function \texttt{plot\_dynamics()} plots a single simulation run, as
returned from the \texttt{run\_simulation()} function. This function is
only provided as a convenience function to provide a way to easily see
the results of a simulation run. An example plot resulting from this
function is shown in \ref{fig:plot-dynamics}.

@@

\caption{Results of a simulation run shown using the function $plot_dynamics()$. In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example. CB  extunderscore1 = cyanobacteria strain 1; SB extunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show the figure produce by $plot_dynamics()$, which produces figures not with all features ideal for publication.}\label{fig:plot-dynamics}
\end{figure}

@@

Here we evidence the impact of the \emph{microxanox} package by
describing three use cases and then by describing how the package can be
a starting point for models of other ecosystems. The first two use cases
are described in detail (including the code for reproducing them) in the
\emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
third is taken from \citet{REF_NEEDED}.

@@

An example of this functionality is given in the \emph{Partial
Reproduction} vignette, which we briefly show here (Figure
\ref{fig:uc1}. The example is composed of a single simulation, at the
beginning of which the system is in the oxic state with high abundance
of cyanobacteria. Oyxgen diffusivity is then slowly decreased and
eventually, around hour 30'000 the system switches to the anoxic state,
with high abundance of both sulfur bacteria types. The oxygen
diffusivity is then increased and at around hour 38'000 the system
abruptly switches back to the oxic state.

@@

\begin{verbatim}
#> 
#> Attaching package: 'dplyr'
#> The following objects are masked from 'package:stats':
#> 
#>     filter, lag
#> The following objects are masked from 'package:base':
#> 
#>     intersect, setdiff, setequal, union
\end{verbatim}

@@

\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter *a*, the oxygen diffusivity) changes. Plot of the stable states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB   extunderscore1 = cyanobacteria strain 1; SB extunderscore1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot_dynamics()$  function.}\label{fig:uc1}
\end{figure}

@@

The results show that the amount of hysteresis depends greatly on the
combinations of organisms present (Figure \ref(figL:uc2)). For example,
with only the CB (cyanobacteria) present, there was no hysteresis. In
contrast, the presence of both CB and SB (sulfate reducing bacteria) led
to a large amount of hysteresis. (These results are also given in the
\emph{Partial Reproduction} vignette.)

@@

The \emph{microxanox} R package allows the simulation, visualisation,
and analysis of a model of a microbial ecosystem while allowing
variation in the amount of diversity containing in each of the
functional groups of organisms present. It has been used for the
research described in another paper that describes one of the first
investigations of the effects of diversity on ecosystem resilience
\citet{Limberger2022}. In that paper, we show that diversity can have
large and important effects of ecosystem responses, highlighting the
need for models such as ours, with which one can easily manipulate the
amount of biodiversity. The \emph{microxanox} package has also been used
to reproduce the results of the paper that inspired the package
development \citet{Bush2017}.

>>>>

  \begin{abstract}
  \(Microxanox\) is an R package to simulate a three functional group
  system (cyanobacteria, phototrophic sulfur bacteria, and
  sulfate-reducing bacteria) with four chemical substrates (phosphorus,
  oxygen, reduced sulfur, and oxidized sulfur) using a set of ordinary
  differential equations. Simulations can be run individually or over a
  parameter range, to find stable states. The model features multiple
  species per functional group, where the number is only limited by
  computational constraints. The R package is constructed in such a way,
  that the results contain the input parameter used, so that a saved
  results can be loaded again and the simulation be repeated.
  \end{abstract}
    \begin{keyword}
    reproducibility \sep regime shift \sep stable state \sep 
    ordinary differential equations
  \end{keyword}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0483}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4138}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.5379}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\centering
\textbf{Nr.}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Code metadata description}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Please fill in this column}
\end{minipage} \\
\midrule()
\endhead
C1 & Current code version & v0.9.0 \\
C2 & Permanent link to code/repository used for this code version &
\url{https://github.com/UZH-PEG/microxanox} \\
& & \href{https://zenodo.org/record/6624125}{10.5281/zenodo.6624125} \\
C3 & Code Ocean compute capsule & \\
C4 & Legal Code License & MIT \\
C5 & Code versioning system used & git \\
C6 & Software code languages, tools, and services used &
\href{https://cran.r-project.org/index.html}{R} \\
C7 & Compilation requirements, operating environments &
\href{https://cran.r-project.org/index.html}{R (\textgreater= 4.1.0)} \\
& & magrittr \\
& & tibble \\
& & ggplot2 \\
& & patchwork \\
& & grDevices \\
& & stats \\
& & mgcv \\
& & deSolve \\
& & dplyr \\
& & tidyr \\
& & stringr \\
& & multidplyr \\
C8 & If available Link to developer documentation/manual &
\href{https://uzh-peg.r-universe.dev/articles/microxanox/User-guide.html}{Usert
Manual} \\
C9 & Support email for questions & Rainer.Krug@uzh.ch;
Owen.Petchey@ieu.uzh \\
\bottomrule()
\end{longtable}

@@

The framework used when writing this package aimed to maximise
simplicity for the user, and to make it easy to reproduce results (see
the supplement \citep{Petchey2022} to \citet{Limberger2022} for an
example of how this is used). As such, all the parameters needed to run
a simulation or find a stable state (i.e.~the final state of the
ecosystem) are contained in a single object (which can easily be created
using included functions). This parameter object is given to a function
that runs the simulations and returns the results. The returned results
object is identical to the parameter object but with an additional slot
named \texttt{results}, which contains the results of the run. Thus the
returned results object contains the simulation conditions (parameters)
as well as the results, and can be used to run the simulation again.
This promotes reproducibility and makes incremental changes of
individual parameters and re-running the simulations straightforward.

@@

The function \texttt{plot\_dynamics()} plots a single simulation run, as
returned from the \texttt{run\_simulation()} function. This function is
only provided as a convenience function to provide a way to easily see
the results of a simulation run. An example plot resulting from this
function is shown in Figure \ref{fig:plot-dynamics}.

@@

\caption{Results of a simulation run shown using the function $plot\_dynamics()$. In this case, there were three strains per functional group, though strains within functional groups had identical properties in this example. CB\_1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB\_1 = phototrophic sulfur bacteria strain 1.}\label{fig:plot-dynamics}
\end{figure}

@@

Here we evidence the impact of the \emph{microxanox} package by
describing three use cases and then by describing how the package can be
a starting point for models of other ecosystems. The first two use cases
are described in detail (including the code for reproducing them) in the
\emph{User Guide} and the \emph{Partial Reproduction} vignettes. The
third is taken from \citet{Limberger2022} and \citet{Petchey2022}.

@@

An example of this functionality is given in the \emph{Partial
Reproduction} vignette, which we briefly show here (Figure
\ref{fig:uc1}). The example is composed of a single simulation, at the
beginning of which the system is in the oxic state with high abundance
of cyanobacteria. Oyxgen diffusivity is then slowly decreased and
eventually, around hour 30'000 the system switches to the anoxic state,
with high abundance of both sulfur bacteria types. The oxygen
diffusivity is then increased and at around hour 38'000 the system
abruptly switches back to the oxic state.

@@

@@

\caption{The temporal dynamics of the ecosystem model when an environmental condition (here parameter $a$, the oxygen diffusivity) changes. Plot of the stable states of the simulation runs under different oxygen diffusivity. In this simulation there is only one strain in each functional group. CB   extunderscore1 = cyanobacteria strain 1; SB\_1 = sulfur reducing bacteria stain 1; PB = phototrophic sulfur bacteria strain 1. Here we show a figure adapted from the output of the $plot\_dynamics()$  function.}\label{fig:uc1}
\end{figure}

@@

The results show that the amount of hysteresis depends greatly on the
combinations of organisms present (Figure \ref{fig:uc2}). For example,
with only the CB (cyanobacteria) present, there was no hysteresis. In
contrast, the presence of both CB and SB (sulfate reducing bacteria) led
to a large amount of hysteresis. (These results are also given in the
\emph{Partial Reproduction} vignette.)

@@

The \emph{microxanox} R package allows the simulation, visualisation,
and analysis of a model of a microbial ecosystem while allowing
variation in the amount of diversity containing in each of the
functional groups of organisms present. It has been used for the
research described in another paper that describes one of the first
investigations of the effects of diversity on ecosystem resilience
\citet{Limberger2022}. In that paper, we show that diversity can have
large and important effects of ecosystem responses, highlighting the
need for models such as ours, with which one can easily manipulate the
amount of biodiversity. The \emph{microxanox} package has also been used
to reproduce the results of the paper that inspired the package
development \citep{Bush2017}.
